---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# iglu

<!-- badges: start -->
[![Travis build status](https://api.travis-ci.com/irinagain/iglu.svg?branch=master)](https://travis-ci.com/github/irinagain/iglu)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/irinagain/iglu?branch=master&svg=true)](https://ci.appveyor.com/project/irinagain/iglu)
<!-- badges: end -->

iglu: Interpreting data from Continuous Glucose Monitors (CGMs)
====================================================================

The R package 'iglu' provides functions for outputting relevant metrics for data collected from Continuous Glucose Monitors (CGM). For reference, see ["Interpretation of continuous glucose monitoring data: glycemic variability and quality of glycemic control." Rodbard  (2009)](https://www.ncbi.nlm.nih.gov/pubmed/19469679).

iglu comes with two example datasets: example_data_1_subject and example_data_5_subject. Each dataset follows the structure iglu's functions are designed around. Note that the 1 subject data is a subset of the 5 subject data. See the examples below for loading and using the data. 

Installation
------------

```{r, eval = FALSE}
# Plain installation
devtools::install_github("irinagain/iglu") # iglu package

# For installation with vignette
devtools::install_github("irinagain/iglu", build_vignettes = TRUE)
```

Example
-------

```{r}
library(iglu)
data(example_data_1_subject) # Load single subject data
## Plot data

# Use plot on dataframe with time and glucose values for time series plot
plot_glu(example_data_1_subject)

# Summary statistics and some metrics
summary_glu(example_data_1_subject)

in_range_percent(example_data_1_subject)

above_percent(example_data_1_subject, targets = c(80,140,200,250))

j_index(example_data_1_subject)

conga(example_data_1_subject)

# Load multiple subject data
data(example_data_5_subject)

plot_glu(example_data_5_subject, plottype = 'lasagna', datatype = 'average')

below_percent(example_data_5_subject, targets = c(80,170,260))

mage(example_data_5_subject)
```

Shiny Demonstration
-------------------

For a demonstration of the package in a point and click interface, click the link below.

https://stevebroll.shinyapps.io/shinyigludemo/
