---
title: "AGP and Episodes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AGP and Episodes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(iglu)
```
The iglu package includes two single page reports - an ambulatory glucose profile (AGP), and an episode calculation report.

# Ambulatory Glucose Profile (AGP)

The iglu package allows one to generate an Ambulatory Glucose Profile (AGP) report - see [Johnson (2019) "Utilizing the ambulatory glucose profile to standardize and implement continuous glucose monitoring in clinical practice."](https://doi.org/10.1089/dia.2019.0034). Below is an example report for Subject 1, which includes information on data collection period, time spent in standardized glycemic ranges (cutoffs of 54, 70, 180 and 250 mg/dL) displayed as a stacked bar chart, glucose variability as measured by %CV, and visualization of quantiles of the glucose profile across days together with daily glucose views. 

```{r, fig.width=10, fig.height=10, out.width="600px", out.height ="500px"}
agp(example_data_1_subject)
```


## Episode Calculation

The Episode_Calculation function measures the number of hypoglycemia and hyperglycemia that occur more than a certain period.

```{r}
episode_calculation(example_data_1_subject,lv2_hypo=70.0, lv2_hyper= 180.0, dur_length = 15)

```

Users can set certain thresholds for the hypo and hyperglycemia by passing parameters, lv1_hypo, lv2_hypo, hy1_hyper, lv2_hpyer. Here, we are passing 70 for the hypoglycemia threshold level 2 threshold and 180 for the hyperglycemia level 2 threshold. By setting a duration length to 15 minutes (the last parameter), the function will count the number of episodes that glucose values go below or above the thresholds more than 15 minutes. 

In this example, we calculated episodes for subject 1. The average numbers of hypo and hyperglycemia for subject 1 (level 1) are 2.7 and 2.6 respectively. It occurred for 111 minutes and 196 minutes in average. You can also see other values such as the mean average values for hypoglycemia and hyperglycemia, and this function can be used for multiple subjects as well.

## Epicalc_profile function

Visualization of the metrics produced by the `episode_calculation()` function is done with the function `epicalc_profile()`. This function takes the `episode_calculation()` output and displays it as a tables of the episode metrics as well as plots that visualizes the subject's glucose level over time. Like `episode_calculation`, this function is designed to work with one subject data at a time, and the structure of the function output is shown below.

```{r}
epicalc_profile(example_data_1_subject)
```

The function will accept the following six parameters that are passed on to `epicalc_profile()`. Like `episode_calculation()`, there are options to adjust the hypoglycemia and hyperglycemia thresholds. If the hyperglycemia threshold is lower than the hypoglycemia threshold, a warning will be produced.

* "`lv1_hypo`" - First threshold for what below is considered hypoglycemia. Default is 100 and the units are mg/dL
* "`lv2_hypo`" - Second threshold for what below is considered hypoglycemia. Default is 70 and the units are mg/dL
* "`lv1_hyper`" - First threshold for what above is considered hyperglycemia. Default is 120 and the units are mg/dL
* "`lv2_hyper`" - Second threshold for what above is considered hyperglycemia. Default is 160 and the units are mg/dL
* "`color_scheme`" - String corresponding to a color scheme for the plot.Default is dark red for `lv1_hypo`, red for `lv2_hypo`, green for normal, yellow for `lv1_hyper`, and orange for `lv2_hyper`.

```{r}
epicalc_profile(example_data_1_subject, lv1_hypo=100, lv2_hypo=70, lv1_hyper=120, lv2_hyper=160, color_scheme="Color Scheme 1")
```
