---
title: "Working in iglu"
author: "Steve Broll"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(iglu)
# date: "`r Sys.Date()`"
```

The iglu package is designed with the following three components of Continuous Glucose Monitor (CGM) data in mind:

* Blood glucose level

* Timestamp for measurement

* Subject identification

There are many metrics used to assess glycemic variability, many of which can be found in ["Interpretation of continuous glucose monitoring data: glycemic variability and quality of glycemic control." Rodbard (2009)](https://www.ncbi.nlm.nih.gov/pubmed/19469679). The iglu package streamlines the calculation of metrics by providing clearly named functions that output metrics with one line of code. 

# Single Subject Workflow
## Data

Example data for 1 subject can be loaded with:

```{r}
data("example_data_1_subject")
```
  
This dataset contains 2915 observations of 3 columns corresponding to the three components listed in the introduction:

* `"id"` - Factor (character string) column for subject identification
* `"time"` - Factor (character string) column that can be converted to DateTime for measurement timestamp
* `"gl"` - Numeric column for glucose measurement


```{r}
dim(example_data_1_subject)
str(example_data_1_subject)
head(example_data_1_subject)
```

Observe that the timestamps in the first rows are not evenly spaced. Linear interpolation is used where appropriate for computation of metrics that require a time component. Time intervals within large gaps between variables will not be interpolated, but instead will be treated as missing values.

## Calculating Metrics

Some metric functions, like `above_percent()`, will return multiple values for a single subject. 

```{r}
above_percent(example_data_1_subject)
```

Subject id will always be printed as a rowname by default, and metrics will be printed as column name. The list of target values for the above_percent metric is a parameter that can be changed:

```{r}
above_percent(example_data_1_subject, targets = c(100, 200, 300))
```

Many metrics have parameters that can be changed. To see available parameters for a given metric, see the documentation i.e. ?above_percent or help(above_percent).  
Some metric functions, like `conga()` (Continuous Overlapping Net Glycemic Action), will return just a single value for each subject.

```{r REMOVE WHEN FIXED, include = F}
library(lubridate); library(intervals)
```

```{r}
conga(example_data_1_subject)
```

Note: `base::as.numeric()` can be used to convert output to a numeric value or vector instead of a dataframe object *only* when working with a single subject.

```{r}
as.numeric(conga(example_data_1_subject))

as.numeric(above_percent(example_data_1_subject))
```

## Plots

When working with a single subject, iglu currently supports the follwing plots: 

* Histogram of glucose values
* Time Series plot of glucose values

The function `plot_glu()` can output either plot, dependent on the input.
For a histogram, simply pass the glucose column of the dataset (numeric).

```{r}
plot_glu(example_data_1_subject$gl) # extract 'gl' column
```

For a time series plot, pass the whole dataset.

```{r}
plot_glu(example_data_1_subject)
```

