---
title: "Working in iglu"
author: "Steve Broll"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(iglu)
# date: "`r Sys.Date()`"
```

The iglu package is designed with the following three components of Continuous Glucose Monitor (CGM) data in mind:

* Blood glucose level

* Timestamp for measurement

* Subject identification

There are many metrics used to assess glycemic variability, many of which can be found in ["Interpretation of continuous glucose monitoring data: glycemic variability and quality of glycemic control." Rodbard (2009)](https://www.ncbi.nlm.nih.gov/pubmed/19469679). The iglu package streamlines the calculation of metrics by providing clearly named functions that output metrics with one line of code. 

# Single Subject Workflow
## Data

Example data for 1 subject can be loaded with:

```{r}
data("example_data_1_subject")
```
  
This dataset contains 2915 observations of 3 columns corresponding to the three components listed in the introduction:

* `"id"` - Factor (character string) column for subject identification
* `"time"` - Factor (character string) column that can be converted to DateTime for measurement timestamp
* `"gl"` - Numeric column for glucose measurement

Data used with iglu functions may have additional columns, but the columns for id, time and glucose values must be named as above.

```{r}
dim(example_data_1_subject)
str(example_data_1_subject)
head(example_data_1_subject)
```

Observe that the timestamps in the first rows are not evenly spaced. Linear interpolation is used where appropriate for computation of metrics that require a time component. Time intervals within large gaps between variables will not be interpolated, but instead will be treated as missing values.

## Calculating Metrics

Some metric functions, like `above_percent()`, will return multiple values for a single subject. 

```{r}
above_percent(example_data_1_subject)
```

Subject id will always be printed as a rowname by default, and metrics will be printed as column name. The list of target values for the above_percent metric is a parameter that can be changed:

```{r}
above_percent(example_data_1_subject, targets = c(100, 200, 300))
```

Many metrics have parameters that can be changed. To see available parameters for a given metric, see the documentation i.e. ?above_percent or help(above_percent).  
Some metric functions, like `conga()` (Continuous Overlapping Net Glycemic Action), will return just a single value for each subject.

```{r}
conga(example_data_1_subject)
```

Note: `base::as.numeric()` can be used to convert output to a numeric value or vector instead of a dataframe object *only* when working with a single subject.

```{r}
as.numeric(conga(example_data_1_subject))

as.numeric(above_percent(example_data_1_subject))
```

# Multiple Subject Workflow

## Data

Example data for multiple subjects can be loaded with:

```{r}
data("example_data_5_subject")
```
  
This dataset contains the same 3 columns as the single subject case, but now with 13866 observations from 5 subjects.

```{r}
dim(example_data_5_subject)
str(example_data_5_subject)
```

## Calculating Metrics

Some metric functions, like `quantile_glu()`, will return multiple values for each subject, producing a dataframe with >1 columns and >1 rows. 

```{r}
quantile_glu(example_data_5_subject)
```

Other metric functions, like `MAGE()` (Mean Amplitude of Glycemic Excursions), will return just a single value for each subject, producing a 1 column dataframe with a row for each subject.

```{r}
conga(example_data_5_subject)
```

Note: `base::as.numeric()` will produce an error when wrapped around a data.frame with more >1 row.

# Visualizations

The iglu package currently supports 3 plot types:

* Time series plot
* Unsorted lasagna plot
* Row sorted lasagna plot

The time series plot can be used on a single subject or multiple subjects, but, in the case of multiple subjects, only the first will be used.
```{r}
plot_glu(example_data_1_subject, plottype = 'tsplot')
plot_glu(example_data_5_subject, plottype = 'tsplot')
```
However, the 'subjects' parameter can be used to select a desired subject.

```{r}
plot_glu(example_data_5_subject, plottype = 'tsplot', subjects = 'Subject 3')
```
The red lines can be shifted to any hypo- and hyperglycemia cutoffs with the 'lower' and 'upper' arguments.

```{r}
plot_glu(example_data_5_subject, plottype = 'tsplot', subjects = 'Subject 3', lower = 80, upper = 150)
```

The unsorted and row sorted lasagna plots are identical in their usage, other than changing the plottype parameter.

```{r}
plot_glu(example_data_5_subject, plottype = 'unsorted')
plot_glu(example_data_5_subject, plottype = 'rowsorted')
```

The 'unsorted' lasagna plot compares the 5 subjects' blood glucose measurements, matching their values in the order the measurements were taken in.  
  
The 'row sorted' lasagna plot sorts the blood glucose measurements into an increasing order.

# Shiny App

The iglu package comes with a shiny app containing all of the metric calculations as well as all plot types of the package itself.
  
The full app can be accessed at [https://stevebroll.shinyapps.io/shinyiglu/](https://stevebroll.shinyapps.io/shinyiglu/), or by running `iglu::run_shiny()` (iglu must be installed to use the run_shiny function).
  
The app itself has a demo available at [https://stevebroll.shinyapps.io/shinyiglu/](https://stevebroll.shinyapps.io/shinyigludemo/) with data pre-loaded.
