---
title: "Getting Started in iglu"
author: "Steve Broll, Elizabeth Chun, Irina Gaynanova"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started in iglu}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(iglu)
# date: "`r Sys.Date()`"
```

The iglu package is designed to assist in the analyses of data from Continuous Glucose Monitors (CGMs). CGMs are small wearable devices that measure glucose levels continuously throughout the day, with some monitors taking measurements as often as every 5 minutes. Data from these monitors provide a detailed quantification of the variation in blood glucose levels during the course of the day, and thus CGMs play an increasing role in clinical practice. For more on CGMs, see [Rodbard (2016) "Continuous Glucose Monitoring: A Review of Successes, Challenges, and Opportunities."](https://doi.org/10.1089/dia.2015.0417).

Multiple CGM-derived metrics have been developed to assess the quality of glycemic control and glycemic variability, many of which are summarized in [Rodbard (2009) "Interpretation of continuous glucose monitoring data: glycemic variability and quality of glycemic control."](https://doi.org/10.1089/dia.2008.0132). The iglu package streamlines the calculation of these metrics by providing clearly named functions that output metrics values with one line of code. 

The iglu package is designed to work with Continuous Glucose Monitor (CGM) data in the form of a data frame with the following three columns present:

* Blood glucose level measurement [in mg/dL] (`"gl"`)

* Timestamp for glucose measurement (`"time"`)

* Subject identification (`"id"`)

The iglu package comes with example data from 5 subjects with Type II diabetes whose glucose levels were measured using a Dexcom G4 CGM. These data are part of a larger study analyzed in [Gaynanova et al. (2020)](https://doi.org/10.1093/biostatistics/kxaa023).

# Shiny App

The iglu package comes with a shiny app containing all of the metric calculations as well as all plot types of the package itself.
  
The full app can be accessed by running `iglu::iglu_shiny()` (iglu must be installed to use the `iglu_shiny` function).
  
The app itself has a demo (reduced functionality) available at [https://stevebroll.shinyapps.io/shinyigludemo/](https://stevebroll.shinyapps.io/shinyigludemo/) with data pre-loaded.

# Example Datasets
The iglu package includes two datasets: example_data_1_subject, and example_data_5_subject. The one subject data is simply the first subject of the five subject data. 

## Example data for 1 subject

Example data with 1 subject can be loaded with:

```{r}
data("example_data_1_subject")
```
  
This dataset contains 2915 observations of 3 columns corresponding to the three components listed in the introduction:

* `"id"` - Factor (character string) column for subject identification
* `"time"` - POSIXct column for datetime values
* `"gl"` - Numeric column for glucose measurement

Data used with iglu functions may have additional columns, but the columns for id, time and glucose values must be named as above.

```{r}
dim(example_data_1_subject)
str(example_data_1_subject)
head(example_data_1_subject)
```

## Example data for 5 subjects

Example data with multiple subjects can be loaded with:

```{r}
data("example_data_5_subject")
```
  
This dataset contains the same 3 columns as the dataset in the single subject case, but now with 13866 observations from 5 subjects. The first subject in this multiple subject dataset is the same as the single subject from the previous examples.

```{r}
dim(example_data_5_subject)
str(example_data_5_subject)
```

# Calculating time-independent metrics

All the metrics implemented in the package can be divided into two categories: time-independent and time-dependent.

Time-independent metrics do not use any linear interpolation because the time component of the data is not used in their calculations. Because the time component is not necessary, when working with a single subject only a glucose vector is required. If a glucose vector for multiple subjects is supplied, or if a data frame that doesn't have all three columns is supplied, these functions will treat all glucose values as though they are from the same subject.

## Example metrics: time-independent

All metric functions in iglu will produce the output in a tibble form. See documentation on tibbles with vignette('tibble') or ?`tbl_df-class`.

Some metric functions, like `above_percent()`, will return multiple values for a single subject. 

```{r}
above_percent(example_data_1_subject)
```

When a data frame is passed, subject id will always be printed in the id column, and metrics will be printed in the following columns. 

As discussed above, just the glucose vector can be supplied for the single subject case. 

```{r}
above_percent(example_data_1_subject$gl)
```

However, it is not recommended to pass just glucose values whenever the time and subject are also available, because this output will not contain the subject ID. 

The list of target values for the above_percent metric is a parameter that can be changed:

```{r}
above_percent(example_data_1_subject, targets = c(100, 200, 300))
```

Many metrics have parameters that can be changed. To see available parameters for a given metric, see the documentation i.e. ?above_percent or help(above_percent).  

Not all metric functions return multiple values. Many, like the Hyperglycemia index metric (function call: `hyper_index()`) will return just a single value for each subject, producing a column for value and a column for subject id (if a dataframe is passed), with one row for each subject.

```{r}
hyper_index(example_data_5_subject)
```

In this example, Subject 2 has the largest Hyperglycemia index, indicating the worst hyperglycemia. This is reflected in percent of times Subject 2 spends above fixed glucose target (see results of `above_percent`). 

# Calculating time-dependent metrics
Unlike time-independent metrics, time-dependent metrics require the input to be a dataframe with subject id's, time, and glucose values. Time-dependent metric functions cannot be passed a vector of glucose values. With timestamped data, a potential challenge arises when timestamps are not on perfect intervals due to missing measurements. To address this challenge, we developed the `CGMS2DayByDay` function.

## CGMS2DayByDay()
Observe that the timestamps in the first rows are not even. The `CGMS2DayByDay` function addresses this issue by linearly interpolating glucose measures for each subject on an equally spaced time grid from day to day. To prevent extrapolation, missing values are inserted between any two measurements that are more than `intergap` minutes apart (default value is 45 minutes, can be changed by the user). This function is automatically called by all metrics that require such interpolation, however it is also available to the user directly. The function is designed to work with one subject data at a time, the structure of function output is shown below.
```{r}
str(CGMS2DayByDay(example_data_1_subject))
```

The first part of the output, `gd2d`, is the interpolated grid of values. Each row corresponds to one day of measurements, and the columns correspond to an equi-distant time grid covering a 24 hour time span. The grid is chosen to match the frequency of the sensor (5 minutes in this example leading to $(24 * 60)/ 5 = 288$ columns), which is returned as `dt0`. The returned `actual_dates` allows one to map the rows in `gd2d` back to the original dates. The achieved alignment of glucose measurement times across the days enables both the calculation of corresponding metrics, and the creation time-dependent visuals such as lasagna plots. The default frequency can be adjusted as follows.

```{r}
str(CGMS2DayByDay(example_data_1_subject, dt0 = 10))
```
Note that the final part of the output reflects our input, and there are now only 144 columns instead of 288.

The CGMS2DayByDay function also allows specification of the maximum allowable gap to interpolate values across (default is 45 minutes) and a string corresponding to time zone (default is the timezone of the user's system). 

Functions for metrics requiring linear interpolation will accept the following three parameters that are passed on to `CGMS2DayByDay()`:

* "`dt0`" - Time frequency (numeric) for interpolation. Default will automatically match the frequency of the data
* "`inter_gap`" - Maximum allowable gap in minutes (numeric) for interpolation
* "`tz`" - String corresponding to timezone where the data's measurements were recorded
  
In the example_data_5_subject dataset, it is important to specify `tz` = 'EST', because a Daylight Savings Time shift can cause miscalculations if the wrong timezone is used. A proper call for this dataset, being recorded in EST, would be: 

```{r, eval = F}
CGMS2DayByDay(example_data_1_subject, tz = 'EST')
```

Examples of proper metric function calls will be shown in the next section.

## Example metrics: time-dependent

Some metric functions, like `conga()` (Continuous Overlapping Net Glycemic Action), will return just a single value for each subject,
resulting in a 2 column tibble (1 column for id and 1 for the single value).

```{r, cache = T}
conga(example_data_1_subject, tz = 'EST')
```

Other metrics can return multiple values for a single subject. For example, `sd_measures()`, which requires linear interpolation, computes 6 unique standard deviation subtypes per subject.

```{r, cache = T}
sd_measures(example_data_5_subject)
```

Notice the high fluctuations in Subject 5, with all but one subtypes of standard deviation being the largest for Subject 5. This provides additional level of CGM data interpretation, since frequent or large glucose fluctuations may contribute to diabetes-related complications independently from chronic hyperglycemia.

## Episode Calculation

The Episode_Calculation function measures the number of hypoglycemia and hyperglycemia that occur more than a certain period.

```{r}
episode_calculation(example_data_5_subject,lv1_hypo=100.0,lv2_hypo = 70, lv1_hyper= 120.0, lv2_hyper = 160, dur_length = 15)
```

Users can set certain thresholds for the hypo and hyperglycemia by passing parameters. Here, we are passing 100 for the hypoglycemia threshold and 120 for the hyperglycemia threshold. By setting a duration length to 15 minutes (the last parameter), the function will count the number of episodes that glucose values go below or above the thresholds more than 15 minutes. 

In this example, we calculated episodes for multiple subjects. The average numbers of hypo and hyperglycemia for subject 1 are 2.7 and 2.6 respectively. It occurred for 111 minutes and 196 minutes in average. You can also see other values such as the mean value of glucose, hypoglycemia and hyperglycemia, and this function can be used for a single subject as well.

## Epicalc_profile function

Visualization of the metrics produced by the `episode_calculation()` function is done with the function `epicalc_profile()`. This function takes the `episode_calculation()` output and displays it as a tables of the episode metrics as well as plots that visualizes the subject's glucose level over time. Like `episode_calculation`, this function is designed to work with one subject data at a time, and the structure of the function output is shown below.
```{r, eval = F}
epicalc_profile(example_data_1_subject)
```
The function will accept the following three parameters that are passed on to `epicalc_profile()`. Like `episode_calculation()`, there are options to adjust the hypoglycemia and hyperglycemia thresholds. If the hyperglycemia threshold is lower than the hypoglycemia threshold, a warning will be produced.
* "`hypo_thres`" - Threshold for what below is considered hypoglycemia. Default is 90 and the units are mg/dL
* "`hyper_thres`" - Threshold for what above is considered hyperglycemia. Default is 120 and the units are mg/dL
* "`color_scheme = c("blue-red", "red-orange")`" - String corresponding to a color scheme for the plot. Default is blue-red
The 'red-orange' color scheme corresponds to red for hypoglycemia, orange for hyperglycemia, and green for in between. The 'blue-red' corresponds to blue for hypoglycemia, red for hyperglycemia, and white for in between.
```{r, eval = F}
epicalc_profile(example_data_1_subject, lv1_hypo=100, lv2_hypo=70, lv1_hyper=120, lv2_hyper=160, color_scheme="Color Scheme 1")


The episode calculation function receives 


# Visualizations

The iglu package supports multiple plot types summarized below:  

| Function call | Visualization description | Main parameters|
|:------------|:------------------------------|:----------|
|`plot_glu` | Multiple plot types: time series and lasagna | `plottype`, `lasagnatype`|
|`plot_roc`| Time series of glucose values colored by rate of change (ROC)| `subjects`, `timelag`|
|`hist_roc` | Histogram of rate of change (ROC) values | `subjects`, `timelag`|
|`plot_lasagna` | Lasagna plot of glucose values for multiple subjects | `datatype`, `lasagnatype`|
|`plot_lasagna_1subject` | Lasagna plot of glucose values for a single subject | `lasagnatype`|
|`agp` | Ambulatory Glucose Profile (AGP) | `maxd`, `daily` |
|`epicalc_profile` | Profile of glycemic episodes | `hypo_thresh`, `hyper_thresh` |
|`mage` | MAGE plot displaying peaks and nadirs | `plot`, `title` |

Time-series and rate of change plots are shown in examples below. For details on the other plotting types, see the specific vignette corresponding to the visualization.

## Example plots: time-series

The simplest visual is the time series plot generated using the function `plot_glu`. This plot type can support both single and multiple subjects.

```{r, fig.width = 7}
plot_glu(example_data_1_subject, plottype = 'tsplot', tz = "EST")
plot_glu(example_data_5_subject, plottype = 'tsplot', tz = "EST")
```

We set the 'tz' (timezone) parameter to be EST because the data was collected in the eastern time zone. If left blank, the time zone used for plotting will be the system's time zone. Time zone is mainly an issue in cases where daylight savings time might make it appear as though there were duplicate values at some time points.

To just plot a single subject of interest from the grid of time series plots, set the 'subjects' parameter to be that subject's ID.

```{r, fig.width = 7}
plot_glu(example_data_5_subject, plottype = 'tsplot', subjects = 'Subject 3', tz = "EST")
```

The red lines can be shifted to any Lower and Upper Target Range Limits with the 'LLTR' and 'ULTR' arguments.

```{r, fig.width = 7}
plot_glu(example_data_5_subject, plottype = 'tsplot', subjects = 'Subject 3', LLTR = 80, ULTR = 150, tz = "EST")
```

## Example plots: Rate of change 

In addition , `iglu` also allows one to visualize local changes in glucose variability as measured by rate of change [Clarke et al. (2009)](https://doi.org/10.1089/dia.2008.0138). There are two types of visualizations associated with rate of change, a time-series plot and a histogram. For both plots, the colors indicate rate of change: white indicates a stable rate of change while red and blue represent times at which the glucose is significantly rising or falling, respectively. Thus colored points represent times of glucose variability, while white points represent glucose stability. The below figure shows a side by side comparison of rate of change time-series plots for two subjects. Subject 1 shows significantly less glucose variability than Subject 5. 

```{r, cache = T, fig.width = 7}
plot_roc(example_data_5_subject, subjects = c("Subject 1", "Subject 5"))
```

The next figure  shows a side by side comparison of rate of change histogram plots for the same subjects. Once again, the colors show in what direction and how quickly the glucose is changing. The histogram plots allow one to immediately assess the variation in rate of change. Extreme values on either end of the histogram indicate very rapid rises or drops in glucose - a high degree of local variability. Here, Subject 1 once again shows lower glucose variability by having a narrower histogram with most values falling between -2 mg/dl/min and 2 mg/dl/min. Subject 5 has a shorter, more widely distributed histogram indicating greater glucose variability.

```{r, cache = T, fig.width = 7}
hist_roc(example_data_5_subject, subjects = c("Subject 1", "Subject 5"))
```


