---
title: "List of metrics implemented in iglu"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{List of metrics implemented in iglu}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(iglu)
```

|Metric Name | Corresponding function in iglu | Reference| Time-dependent |
|:------------|:------------------------------:|:----------|:------------|
|Active Percent | active_percent | [Danne et al. (2017)](https://doi.org/10.2337/dc17-1600)| Yes |
|ADRR (Average Daily Risk Range) | adrr | [Kovatchev et al. (2006)](https://doi.org/10.2337/dc06-1085)| Yes |
|AUC (Area Under the Curve) | auc | [Danne et al. (2017)](https://doi.org/10.2337/dc17-1600)| Yes |
|CONGA (Continuous Overall Net Glycemic Action) | conga | [McDonnell et al. (2005)](https://doi.org/10.1089/dia.2005.7.253)| Yes |
|Coefficient of variation (CV) | cv_glu | [Rodbard (2009)](https://doi.org/10.1089/dia.2008.0132) | No |
|CVmean (mean of daily CV) | cv_measures (CVmean) | [Umpierrez et al. (2018)](https://doi.org/10.1016/j.amjms.2018.09.010) | Yes |
|CVsd (SD of daily CV) | cv_measures (CVsd) | [Umpierrez et al. (2018)](https://doi.org/10.1016/j.amjms.2018.09.010) | Yes |
|eA1c (Estimated A1c) | ea1c | [Nathan et. al (2008)](https://doi.org/10.2337/dc08-0545)| No |
|Glucose Management Indicator (GMI) | gmi | [Bergenstal (2018)](https://doi.org/10.2337/dc18-1581)| No |
|GRADE (Glycemic Risk Assessment Diabetes Equation) | grade | [Hill et al. (2007)](https://doi.org/10.1111/j.1464-5491.2007.02119.x)| No |
|GRADEeu | grade_eugly | [Hill et al. (2007)](https://doi.org/10.1111/j.1464-5491.2007.02119.x)| No |
|GRADEhyper | grade_hyper | [Hill et al. (2007)](https://doi.org/10.1111/j.1464-5491.2007.02119.x)| No |
|GRADEhypo | grade_hypo | [Hill et al. (2007)](https://doi.org/10.1111/j.1464-5491.2007.02119.x)| No |
|GVP (Glucose Variability Percentage) | gvp | [Peyser et al. (2017)](https://dx.doi.org/10.1089%2Fdia.2017.0187) | Yes |
|HBGI (High Blood Glucose Index) | hbgi | [Kovatchev et al. (2006)](https://doi.org/10.2337/dc06-1085)| No |
|LBGI (Low Blood Glucose Index) | lbgi | [Kovatchev et al. (2006)](https://doi.org/10.2337/dc06-1085)| No |
|Hyperglycemia Index Parameters | hyper_index | [Rodbard (2009)](https://doi.org/10.1089/dia.2008.0132)| No |
|Hypoglycemia Index Parameters | hypo_index | [Rodbard (2009)](https://doi.org/10.1089/dia.2008.0132)| No |
|IGC (Index of Glycemic Control)| igc | [Rodbard (2009)](https://doi.org/10.1089/dia.2008.0132) | No |
|IQR (Interquartile Range)| iqr_glu | | No |
|J-Index| j_index | [Wojcicki (1995)](https://doi.org/10.1055/s-2007-979906) | No |
|M-value| m_value | [Schlichtkrull et al. (1965)](https://doi.org/10.1111/j.0954-6820.1965.tb01810.x) |  No |
|MAD (Mean Absolute Deviation) | mad_glu | | No |
|MAGE (Mean Amplitude of Glycemic Excursions)| mage | [Service et al. (1970)](https://doi.org/10.2337/diab.19.9.644) | No |
|Mean   | mean_glu | |  No |
|Median | median_glu | | No |
|MODD (Mean Of Daily Differences)| modd | [Nelson (1980)](https://doi.org/10.2337/diacare.3.1.58) | Yes |
|Percent Above | above_percent | | No |
|Percent Below | below_percent | | No |
|Percent in range | in_range_percent | | No |
|Quantiles | quantile_glu | | No |
|Range | range_glu | | No |
|Rate of change (ROC) | roc | [Clarke et al. (2009)](https://doi.org/10.1089/dia.2008.0138) | Yes |
|Standard Deviation of ROC | sd_roc | [Clarke et al. (2009)](https://doi.org/10.1089/dia.2008.0138) | Yes |
|Standard Deviation | sd_glu | | No |
|SdB (between days, within timepoints)| sd_measures (SdB) |[Rodbard (2009)](https://doi.org/10.1089/dia.2009.0015)| Yes |
|SdBDM (between days, within timepoints, corrected for changes in daily means)| sd_measures (SdBDM) |[Rodbard (2009)](https://doi.org/10.1089/dia.2009.0015)| Yes |
|SdDM (horizontal sd)| sd_measures (SdDM) |[Rodbard (2009)](https://doi.org/10.1089/dia.2009.0015)| Yes |
|SdHHMM (between time points)| sd_measures (SdHHMM) |[Rodbard (2009)](https://doi.org/10.1089/dia.2009.0015)| Yes |
|SdW (vertical within days)| sd_measures (Sd)|[Rodbard (2009)](https://doi.org/10.1089/dia.2009.0015)| Yes |
|SdWSH (within time series)| sd_measures (SdWSH) |[Rodbard (2009)](https://doi.org/10.1089/dia.2009.0015)| Yes |
