---
title: "List of metrics implemented in iglu"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{List of metrics implemented in iglu}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(iglu)
```

|Metric Name | Corresponding function in iglu | Reference|
|------------|:------------------------------:|----------|
|ADRR (Average Daily Risk Range) | adrr |[Kovatchev et al. (2006)](https://doi.org/10.2337/dc06-1085)|
|CONGA (continuous overall net glycemic action) | conga | [McDonnell et al. (2005)](https://doi.org/10.1089/dia.2005.7.253)|
|GRADE (Glycemic Risk Assessment Diabetes Equation) | grade | [Hill et al. (2007)](https://doi.org/10.1111/j.1464-5491.2007.02119.x)|
|GRADEeu | grade_eugly | [Hill et al. (2007)](https://doi.org/10.1111/j.1464-5491.2007.02119.x)|
|GRADEhyper | grade_hyper | [Hill et al. (2007)](https://doi.org/10.1111/j.1464-5491.2007.02119.x)|
|GRADEhypo | grade_hypo | [Hill et al. (2007)](https://doi.org/10.1111/j.1464-5491.2007.02119.x)|
|HBGI (High Blood Glucose Index) | hbgi | [Kovatchev et al. (2006)](https://doi.org/10.2337/dc06-1085)|
|LBGI (Low Blood Glucose Index) | lbgi | [Kovatchev et al. (2006)](https://doi.org/10.2337/dc06-1085)|
|Hyperglycemia Index parameters | hyper_index | [Rodbard (2009)](https://doi.org/10.1089/dia.2008.0132)|
|Hypoglycemia Index parameters | hypo_index | [Rodbard (2009)](https://doi.org/10.1089/dia.2008.0132)|
|IGC (Index of Glycemic Control)| igc | [Rodbard (2009)](https://doi.org/10.1089/dia.2008.0132) |
|IQR (Interquartile Range)| iqr_glu | |
|J-Index| j_index | [Wojcicki (1995)](https://doi.org/10.1055/s-2007-979906) |
|M-value| m_value | [Schlichtkrull et al. (1965)](https://doi.org/10.1111/j.0954-6820.1965.tb01810.x)
|MAGE (Mean Amplitude of Glycemic Excursions)| mage | [Service et al. (1970)](https://doi.org/10.2337/diab.19.9.644) |
|Mean        | mean_glu                       |          |
|MODD (Mean Of Daily Differences)| modd | [Nelson (1980)](https://doi.org/10.2337/diacare.3.1.58) |
|SD | sd_glu | |
|SdB (between days, within timepoints)| sd_measures (SdB) |[Rodbard (2009)](https://doi.org/10.1089/dia.2009.0015)|
|SdBDM (between days, within timepoints, corrected for changes in daily means)| sd_measures (SdBDM) |[Rodbard (2009)](https://doi.org/10.1089/dia.2009.0015)|
|SdDM (horizontal sd)| sd_measures (SdDM) |[Rodbard (2009)](https://doi.org/10.1089/dia.2009.0015)|
|SdHHMM (between time points)| sd_measures (SdHHMM) |[Rodbard (2009)](https://doi.org/10.1089/dia.2009.0015)|
|SdW (vertical within days)| sd_measures (Sd)|[Rodbard (2009)](https://doi.org/10.1089/dia.2009.0015)|
|SdWSH (within time series)| sd_measures (SdWSH) |[Rodbard (2009)](https://doi.org/10.1089/dia.2009.0015)|
